A new tidy data structure to support exploration and modeling of temporal data

### 2019-08-29 tsibble.Rmd
Improved wording

Temporal data problems often fall into two types of analysis, time series and longitudinal. Both of these may have very similar data input, but the representation for modelling is typically different. Time series analysis tends to focus on the dependency with series, and the cross-correlation between series. Longitudinal analysis tends to focus on overall temporal patterns across demographic or experimental treatment strata, that incorporates within subject dependency.

@@

Implicit missingness are values that should be present but are absent. In regularly-spaced temporal data, these are data entries that should be available at certain time points but are missing, leaving gaps in time. These can be detected when computing the interval estimate. It will be a problem for temporal models and operations like lag/lead are applied. A family of verbs are provided to help explore implicit missing values, and convert them into an explicit state, as follows:

@@

A common practice with energy data analysis is load forecasting, because providers need to know they have capacity to supply electricity. To illustrate the pipeline including modeling, here demand is modeled for December 2013, with the usage forecast for the last day (48 steps ahead because the data is half-hourly). The energy data for the last day is not used for modeling. ARIMA models with and without temperature information are fitted using automatic order selection [@autoarima]. The logarithmic transformation is applied to the average demand to ensure the positive forecasts. Figure \@ref(fig:elec-forecast) plots the one-day forecasts from both models against the actual demand, for the last two-week window. ARIMA regressed by average temperatures gives a better fit than the one without, although both tend to underestimate the night demand. The forecasting performance is reported in Table \@ref(tab:elec-accuracy), consistent with the findings in Figure \@ref(fig:elec-forecast).

@@

(ref:elec-forecast-cap) One-day (48 steps ahead) forecasts generated by ARIMA models, with and without temperatures, plotted against the actual demand. Both nicely capture the temporal dynamics, but ARIMA with temperature performs better than the model without.

>>>>

Temporal data problems often fall into two types of analysis, time series and longitudinal. Both of these may have similar data input, but the representation for modelling is typically different. Time series analysis tends to focus on the dependency with series, and the cross-correlation between series. Longitudinal analysis tends to focus on overall temporal patterns across demographic or experimental treatment strata, that incorporates within subject dependency.

@@

Implicit missingness are values that should be present but are absent. In regularly spaced temporal data, these are data entries that should be available at certain time points but are missing, leaving gaps in time. These can be detected when computing the interval estimate. It will be a problem for temporal models and operations like lag/lead are applied. A family of verbs are provided to help explore implicit missing values, and convert them into an explicit state, as follows:

@@

A common practice with energy data analysis is load forecasting, because providers need to know they have capacity to supply electricity. To illustrate the pipeline including modeling, here demand is modeled for December 2013, with the usage forecast for the last day (48 steps ahead because the data is half-hourly). The energy data for the last day is not used for modeling. ARIMA models with and without a temperature covariate are fitted using automatic order selection [@autoarima]. The logarithmic transformation is applied to the average demand to ensure the positive forecasts. Figure \@ref(fig:elec-forecast) plots the one-day forecasts from both models against the actual demand, for the last two-week window. ARIMA regressed by average temperatures gives a better fit than the one without, although both tend to underestimate the night demand. The forecasting performance is reported in Table \@ref(tab:elec-accuracy), consistent with the findings in Figure \@ref(fig:elec-forecast).

@@

(ref:elec-forecast-cap) One-day (48 steps ahead) forecasts generated by ARIMA models, with and without a temperature covariate, plotted against the actual demand. Both nicely capture the temporal dynamics, but ARIMA with temperature performs better than the model without.
